---
title: "GDSC - Fashion Trend Analysis Project Report"
date: "`r Sys.Date()`"
output: pdf_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

To investigate the relationship between trends within different sectors
of fashion, and see how they influence one another, we collected data
and performed some experiments to test some of our hypotheses. We
started by manually collecting data from a variety of sources, and
creating CSV files that could be used in our report. We acknowledge that
manual data collection involves some human error, but this step was
crucial for learning the entire process of conducting experiments, and
for coming up with appropriate hypotheses. In the future, this data
collection could be automated using web-scraping, pattern recognition,
and an AI/ML model, which is something our team plans to pursue further
in the future.

### Getting Started

The following are the necessary packages we use within our testing:

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
library(tinytex)
library(rcompanion)
```

Next, we create data frames from each CSV file containing separately
collected data, allowing for easy implementation and analysis:

```{r}
depop <- data.frame(read.csv("Depop Data Collection - Sheet1.csv"))

hm <- data.frame(read.csv("Hannah Wen Data Final - Sheet1.csv"))

shein <- data.frame(read.csv("Hannah Wen Data Final - Sheet2.csv"))

highfashion_dresses <- data.frame(read.csv("High Fashion Data Collection - dress trends.csv"))

highfashion_spring <- data.frame(read.csv("High Fashion Data Collection - overall spring trends.csv"))

pinterest <- data.frame(read.csv("Pinterest Data Collection - Sheet1.csv"))

nordstrom <- data.frame(read.csv("Zara_Nordstrom Dataset - nordstrom (1).csv"))

zara <- data.frame(read.csv("Zara_Nordstrom Dataset - zara (1).csv"))
```

## First Experiment:

### Are midi dresses the most on-trend amongst all clothing retailers for current spring/summer season?

Hypothesis: 
Midi dresses are more common than other dress lengths for all retailers.

In order to test this hypothesis we will run a Goodness of Fit test for
each retailer.

First we have to ensure that all requirements are met for the test.

Chi Squared Goodness of Fit Test Requirements:

-   Samples are independent: there is human error because data was
    collected manually but they are all randomly selected from the
    "Trending" or recommended pages of the retailer
-   The data is categorical with the quality we are interested in being
    dress length with levels: mini, midi, maxi -Expected values are
    greater than 5 for each category. We have three categories so
    3\*5=15 so n must be \>= 15

Formatting sets in a way that makes them easier to manipulate later:
```{r}
Retailer = c("Zara")
zara2 <- zara %>%
  rename("Length" = "Length..dress.skirts.")
zaraF <- cbind(zara2, Retailer)%>%
    filter(Article == "dress")

nordstrom2 <- nordstrom %>%
  rename("Length" = "Length..dress.skirts.")
Retailer=c("Nordstrom")
nordstromF <- cbind(nordstrom2, Retailer)%>%
    filter(Article == "dress")

hmF <- hm %>%
  rename("Length" = "Length..dress.skirts.")%>%
  filter(Article == "dress")

sheinF <- shein %>% 
  rename("Length" = "Length..dress.skirts.")

Retailer = c("Depop")
depopF <- cbind(depop, Retailer)%>%
    filter(Article == "dress")

Retailer = c("Pinterest")
pinterestF <- cbind(pinterest, Retailer)%>%
  filter(Article == "dress")
  
highF <- highfashion_dresses %>%
  filter(article == "dress")
```

Creating frequency tables for mini, midi, and maxi length dresses to see
if the necessary sample size is met for each data set:
```{r}
zaraCount <- zaraF %>%
  group_by(Length)%>%
  summarize(count = n())
sum(zaraCount$count)

nordstromCount <- nordstromF%>%
  group_by(Length)%>%
  summarize(count = n())
sum(nordstromCount$count)

depopCount <- depopF %>%
  group_by(Length) %>%
  summarize(count=n())
sum(depopCount$count)

pinCount <- pinterestF %>%
  group_by(Length)%>%
  summarize(count = n())
sum(pinCount$count)

sheinCount <- sheinF%>%
  group_by(Length) %>%
  summarize(count=n())
sum(sheinCount$count)

hmCount <- hmF%>%
  group_by(Length) %>%
  summarize(count=n())
sum(hmCount$count)

highCount<- highfashion_dresses%>%
  group_by(length) %>%
  summarize(count=n())
sum(highCount$count)
```

Zara: n = 17 Nordstrom: n = 8 Depop: n = 16 Pinterest: n = 3 SHEIN: n=11
H&M: n = 21 High Fashion: n = 30

We find that the third requirement of expected value is met for Zara,
Depop, H&M, and high fashion. Unfortunately, the remainder of the
retailers/data do not meet the requirements, and will therefore be left
out of this specific hypothesis testing. Since our data was not
collected to accomodate this number of dresses, a potential point of
growth could be collecting more dress data from those brands.

$H_0$: The proportion of mini, midi, and maxi length dresses are the
same. $H_A$: The proportion of mini, midi, and maxi length dresses are
not the same.

$\alpha$=0.05 (significance level)

The following will plot the distribution of lengths, as well as perform
a chi-squared test for each source of data (retailer).

Zara Dresses:
```{r}
g <- ggplot(zaraF, aes(Length))
g+geom_bar()+ggtitle("Zara Dress Length Distribution")
resZara <-chisq.test(zaraCount$count, p=rep(1/length(zaraCount$Length), 3))
resZara
```

p-value \> $\alpha$

Since the p-value is greater than the alpha level we fail to reject the
null hypothesis and conclude that the proportions of mini, midi, and
maxi length dresses are the same. This means that the variation in the
frequency of dress length in our sample size does not necessarily mean
that there is a variation in frequency in the population.

This disproves our original question of whether midi dresses are the
most popular dress length amongst all retailers, however, now we can
analyze the trends of individual retailers.

Depop Dresses:
```{r}
g <- ggplot(depopF, aes(Length))
g+geom_bar() + ggtitle("Depop Dress Length Distribution")
resDepop <-chisq.test(depopCount$count, p=rep(1/length(depopCount$Length), 3))

resDepop
```

p-value \< $\alpha$

Since the p-value is less than the alpha level we reject the null
hypothesis and conclude that the proportion of mini, midi, and maxi
length dresses are not the same.

We can see in the histogram that there is a larger number of midi-length
dresses than there are mini and maxi which can lead us to infer that
that length is the most popular dress length on Depop in comparison to
the other lengths.

H&M dresses:
```{r}
g <- ggplot(hmF, aes(Length))
g+geom_bar()+ggtitle("H&M Dress Length Distribution")
reshm <-chisq.test(hmCount$count, p=rep(1/length(hmCount$Length), 3))
reshm
```

p-value \< $\alpha$

Since the p-value is less than the alpha level we reject the null
hypothesis and conlude that the proportion of mini, midi, and maxi
length dresses are not the same.

Here we see that there is a larger number of mini dresses so we can
infer that this is the variable that is causing us to reject the null
hypothesis. However, this disproves our original hypothesis that midi
length dresses are the most popular among all retailers as this test
supports that the most popular dress length at H&M is mini.

High Fashion Dresses:
```{r}
g <- ggplot(highF, aes(length))
g+geom_bar()+ggtitle("High Fashion Dress Length Distribution")

resHigh <-chisq.test(highCount$count, p=rep(1/length(highCount$length), 3))
resHigh
```
p-value \< $\alpha$

Since the p-value is less than the alpha level we reject the null
hypothesis and conclude that the proportion of mini, midi, and maxi
length dresses are not the same.

Both maxi and midi dresses have relatively high frequencies in this
distribution so we can infer that those are the variables causing us to
reject the null hypothesis. However, they are relatively even in number
so it is inconclusive whether midi or maxi length dresses is most
popular although we can conculsively infer that mini dresses are the
least popular.

### Conclusion About Dress Length:

Despite having noticed a lot of midi dresses in the process of our data
collection, our results showed that the only retailer where midi dresses
are most on-trend is Depop. This is an interesting observation that can
give us information on how trends are evolving. Depop is a second-hand
retailer that sells primarily to young-adults; this may indicate that
midi dresses are being produced less by modern fashion retailers, but
are potentialy coming back in-style for young adults, as they are being
sold more used than new.

## Second Experiment:

### At what price-point can we find the most skirts and dresses with pockets? What does this say about the cost of comfort in women's clothing?

We looked at different clothing retailers at various prices, ranging
from \$5 pieces at SHEIN and \$5000 pieces from high fashion brands, to
see where we are most likely to find women's skirts and dresses with
pockets.

```{r}
# no dresses or skirts with pockets 
depop %>% filter(Article %in% c('dress', 'skirt')) 
# no dresses or skirts with pockets among h&m or shein 

highfashion_dresses %>% count(pockets) # 3 dresses with pockets out of 30 
highfashion_spring %>% filter(article %in% c('dress', 'skirt')) %>% count(pockets) # 2/11

pinterest %>% filter(Article %in% c('dress', 'skirt')) # no pockets among these in pinterest

nordstrom %>% filter(Article %in% c('dress', 'skirt')) # no pockets

zara %>% filter(Article %in% c('dress', 'skirt')) %>% count(Pockets) # 2 out of 19 
```
Data observations pulled from Depop, Shein, H&M, Nordstrom, and
Pinterest did not have any dresses or skirts with pockets. Dresses
pulled from the high fashion dresses data frame demonstrated a 10%
pocket rate with 3 out of 30 dresses having pockets. Skirts and dresses
pulled from the general high fashion spring trends data frame
demonstrated a 18% pocket rate with 2 out of 11 articles having pockets.
Zara has approximately a 10% pocket rate with 2 articles out of 19
having pockets.

This implies that pockets in skirts and dresses are more common at a
higher price level, while some of the cheapest retailers had zero
dresses or skirts with pockets. However, it is important to observe that
the sample size in each data frame is small, and that the high fashion
data frames pull from different brands while data frames like the Zara
or H&M one pull from only one. In general, it is observed that pockets
are a rare occurrence among skirts and dresses.

### Pockets Conclusion:

Seeing as that pockets are more likely to be found in high-end fashion,
they can somewhat be considered a "luxury," and many women would
probably agree. The proportions were all very low, showing that it is
extremely rare to find dresses or skirts with pockets. Although we only
investigated women's clothing, it would be very interesting to see how
common pockets are in men's clothing in comparison. This future
experiment could reveal a lot about the lack of priority that comfort
has in women's fashion compared to men's fashion, and what this means in
modern society.

## Third Experiment:

### How similar are the materials that are used among different retailers at different price points?

Most retailers sell clothes that consist of synthetic fibers such as
polyester, viscose, and rayon. Garments made of these materials will not
be considered since they are so popular, especially among fast fashion
retailers. While some data frames such as the high fashion data frame
recorded every single material, some did not mention the extremely
common materials. In the SHEIN data frame, no material was recorded
since every garment observed was some combination of the common
materials (polyester, viscose, rayon). Thus, it is excluded from these
tests. A chi-squared test of independence is typically performed when
testing whether two categorical variables are independent. However, do
to the small sample size, we will use the Fisher's exact probability
test.

We need to create proportions for materials, and set up the hypothesis
test.

Data Manipulation:
```{r}
# pulling vectors of materials for each category
depop_materials = depop %>% pull(Material)
hm_materials = hm %>% pull(Material)
# no materials for shein recorded because there was no material observed that was not one of the typical materials
highfashion_dress_materials = highfashion_dresses %>% pull(material) 
highfashion_spring_materials = highfashion_spring %>% pull(material)
pinterest_materials = pinterest %>% pull(Material)
nordstrom_materials = nordstrom %>% pull(Material)
zara_materials = zara %>% pull(Material)
```

```{r}
table(depop_materials)
table(hm_materials)
table(highfashion_dress_materials)
table(highfashion_spring_materials)
table(pinterest_materials)
table(nordstrom_materials)
table(zara_materials)
```

Now that we can find all the counts for each material we want to perform
a test on, we can begin testing. Taking a general look of the contents
of the materials, we will calculate the proportions of the materials:
silk/satin, cotton/linen, and jersey.

Due to the small sample size, Fisher's Exact Test will be performed.
This test requires the assumptions that the data collected is a random
sample, that the data points are dependent of each other, and that the
groups of our categorical variable are mutually exclusive, meaning that
a material cannot be both found and not found in a garment.

The Fisher Exact Test will return a p-value, which is the probability
that the null hypothesis is true. Our null hypothesis for each test will
be that the proportion of a certain material is constant among each type
of retailer. This test will be performed at the commonly used 0.05 level
of significance, meaning that if the p-value is less than 0.05, there is
sufficient evidence to reject the null hypothesis. The alternative
hypothesis is that the proportions are not the same among every
retailer.

Calculations and performing the Fisher test:
```{r}
silk.satin.n = c(1, 2, 5, 1, 1, 0, 0) # count of silk or satin found in each dataframe
n = c(length(depop_materials), length(hm_materials), length(highfashion_dress_materials), 
      length(highfashion_spring_materials), length(pinterest_materials), length(nordstrom_materials),
      length(zara_materials)) # vector of length of each of the material vectors
names = c("Depop", 'H&M', 'High fashion dresses', 'High fashion spring trends',
               'Pinterest', 'Nordstrom', 'Zara') # row names 
silk.satin = data.frame("Silk/Satin" = silk.satin.n, "Silk/satin not found" = n - silk.satin.n, 
                        row.names = names)
silk.satin
fisher.test(silk.satin)
```
The p-value is approximately 0.005, which means that we reject the null
hypothesis that the proportion of satin and/or silk found in garments is
not the same among all retailers.

```{r}
cotton.linen.n = c(6, 2, 12, 5, 3, 11, 22) # vector of count of cotton or linen found
cotton.linen = data.frame("Cotton/Linen" = cotton.linen.n, "Cotton/linen not found" = n - cotton.linen.n, 
                row.names = names) 
cotton.linen
fisher.test(cotton.linen, workspace=2e8)
```
The p-value is extremely close to 0, providing sufficient evidence to
reject the null hypothesis and suggest that the proportions of cotton
and/or linen are not the same among each retailer.

```{r}
jersey.n = c(0, 3, 1, 6, 0, 0, 0) # vector of counts of jersey 
jersey = data.frame("Jersey" = jersey.n, "Jersey not found" = n - jersey.n, 
                row.names = names)
jersey
fisher.test(jersey)
```
The p-value is incredibly close to 0, indicating that we reject the null
hypothesis also.

All 3 tests provide sufficient evidence to reject the null hypotheses,
suggesting that none of these materials are found at the same rate among
every retailer. However, the p-values of jersey and cotton/linen were
much smaller than the p-value for satin/silk, meaning that there is a
higher chance of the same proportions of satin/silk being found in these
retailers again.

### Visualizations to Display Material Trends:

```{r}
X = names = c("Depop", 'H&M', 'HF Dresses', 'HF Spring',
               'Pinterest', 'Nordstrom', 'Zara')
silksat.df = data.frame(X, silk.satin.n, n - silk.satin.n)
silk.sat.df = 
  pivot_longer(silksat.df, cols = !X, names_to = 'Type', values_to = 'Count')
ggplot(silk.sat.df, aes(x = X, y = Count, fill = Type)) + geom_bar(stat = 'identity',
        position = position_stack(reverse=TRUE)) + theme_minimal() + 
  scale_fill_brewer(palette = 'Pastel1',labels = c('Not found', 'Found')) + 
  labs(x = 'Retailer', title = 'Silk and Satin in Garments Among Retailers') +
  theme(legend.position = "none")
```

```{r}
cotton.linen.df = data.frame(X, cotton.linen.n, n - cotton.linen.n)
cotton.linen.df = 
  pivot_longer(cotton.linen.df, cols = !X, names_to = 'Type', values_to = 'Count')
ggplot(cotton.linen.df, aes(x = X, y = Count, fill = Type)) + geom_bar(stat = 'identity',
        position = position_stack(reverse=FALSE)) + theme_minimal() + 
  scale_fill_brewer(palette = 'Pastel2',labels = c('Not found', 'Found')) + 
  labs(x = 'Retailer', title = 'Cotton and Linen in Garments Among Retailers') +
  theme(legend.position = "none")
```

```{r}
jersey.df = data.frame(X, jersey.n, n - jersey.n)
jersey.df = 
  pivot_longer(jersey.df, cols = !X, names_to = 'Type', values_to = 'Count')
ggplot(jersey.df, aes(x = X, y = Count, fill = Type)) + geom_bar(stat = 'identity',
        position = position_stack(reverse=FALSE)) + theme_minimal() + 
  scale_fill_brewer(palette = 'Accent',labels = c('Not found', 'Found')) + 
  labs(x = 'Retailer', title = 'Jersey in Garments Among Retailers') +
  theme(legend.position = "none") 
```

### Conclusion About Common Materials:

We can see from our testing that certain materials are trending within
this spring/summer fashion season, which makes sense in accordance with
the weather. Additionally, analyzing proportions of materials seen in
different retailers at different price-points may also indicate the
higher use of expensive materials in high-fashion compared to cheaper
materials in mainstream consumer fashion.

## Fourth Experiment:

### Is there an association between colors found in high fashion products and colors found in second-hand retail products? If so, to what extent?

Hypothesis: There is a strong positive correlation between colors found
in high fashion products and colors found in second-hand retail
products. In other words, colors that are seen commonly in high fashion
products will also be seen more often in second-hand retail products.

### Process:

Conduct a Fisher's Exact Test to determine association between colors
found in High Fashion products and colors found in second-hand retail
products. If there is an association, run Cramer's V test to measure the
association. (Note: I initially attempted to use a Chi-Squared test, but
the values per cell were too low. Many values were zero or \<=5. For
smaller sample sizes, Fisher's test is more suitable.)

### Manipulating Data Frames:

Combining the high fashion data frames to create a single one:
```{r}
highfashion <- full_join(highfashion_dresses, highfashion_spring) 
highfashion <- highfashion %>% rename("Color" = "color") #all other data sets use the capitalized version as the column name
```

Creating columns indicating the categories of each item:
```{r}
highfashion <- highfashion %>% mutate(Category = "High Fashion")
depop <- depop %>% mutate(Category = "Depop")
nordstrom <- nordstrom %>% mutate(Category = "Nordstrom")
zara <- zara %>% mutate(Category = "Zara")
hm <- hm %>% mutate(Category = "H&M")
shein <- depop %>% mutate(Category = "Shein")
```

### Creating Contingency Tables

Creating a contingency table for the High Fashion and Depop datasets
based on the frequncy of colors by category:
```{r}
high_depop <- full_join(depop, highfashion, by = c("Category", "Color")) #joining the depop and high fashion datasets into a singular one based on the color 
high_depop <- high_depop %>% distinct() #ensuring there are no duplicate entries

table.high_depop <- high_depop %>% select(c("Category", "Color")) %>% table() #creating a contingency table based on only the category and color variables. shows the frequency of each color for each category
table.high_depop
```

### High Fashion and Depop: Tests & Visualizations

Sampling is done at random & outcomes are mutually exclusive

Null hypothesis (H0): There is no association between the two
variables.\
Alternative hypothesis (HA): There is an association between the two
variables.

Fisher's Exact Test:\
p-value of 0.01182\
0.01182 \< alpha = 0.05\
Null hypothesis can be rejected. Therefore, there is an association
between the colors found in High Fashion and Depop products.

Cramer's V:\
Cramer's V measures the association between two variables with a result
of 0 indicating no association and a result of 1 indication a perfect
association. Cramer's V for the association between colors found in High
Fashion and Depop products is 0.5332. This indicates a somewhat strong
association.

Conducting the tests:
```{r}
degrees_of_freedom <- min(nrow(table.high_depop)-1, ncol(table.high_depop)-1)
degrees_of_freedom
fisher.test(high_depop$Category,high_depop$Color)
cramerV(high_depop$Category,high_depop$Color)
```
Note: This visualization (and analysis overall) may be improved by
grouping together certain colors into more general categories. ie:
"off-white" and "white" can be combined into just "white" or "denim" and
"blue" can be combined into just "blue".

Visualizing the distributions by plotting:
```{r}
color.high_depop <- high_depop %>% select(c("Category", "Color"))

ggplot(color.high_depop, aes(x=Color, fill= Category)) + 
  geom_bar(position = "dodge")+
  scale_fill_brewer(palette = 'Paired',labels = c('Depop', 'High Fashion')) + 
  scale_x_discrete(guide = guide_axis(angle = 90))+
  ylab("Count")+
  ggtitle("The Frequency of Items of a Certain Color in Depop and High Fashion")
```

## Fifth Experiment:

### Are trousers becoming more popular than other styles of bottoms?

Hypothesis:
Trousers are seen more in current fashion trends than other types of
bottoms

Process:
Conduct a two-sided z-test to compare the proportion of trousers compared to other styles of bottoms.

Data Manipulation:
```{r}
# filter out only the rows with pants or shorts

pinterest_pants <- pinterest %>% 
  filter(Article %in% c("pants", "shorts"))

hm_pants <- hm %>% # none
  filter(Article %in% c("pants", "shorts"))
  
shein_pants <- shein %>% # none
  filter(Article %in% c("pants", "shorts"))

depop_pants <- depop %>% 
  filter(Article %in% c("pants", "shorts"))
  
highfashion_spring_pants <- highfashion_spring %>% 
  filter(article %in% c("pants", "shorts"))

nordstrom_pants <- nordstrom %>% 
  filter(Article %in% c("pants", "shorts"))

zara_pants <-  zara %>% # none
  filter(Article %in% c("pants", "shorts"))

head(pinterest_pants) # 12 trousers

highfashion_spring_pants <- rename(highfashion_spring_pants, Article = article) 
head(highfashion_spring_pants) # 1 trouser

head(nordstrom_pants) # 1 trouser

```

Data Organization:
```{r}
# make data frame with non-trousers

pinterest_other <- pinterest_pants %>% 
  filter(is.na(Embellishment) | Embellishment != "trousers")

highfashions_other <- highfashion_spring_pants %>% 
  filter(material ==  "denim")

nordstrom_other <- nordstrom_pants %>% 
  filter(Material != "linen/cotton")


other_pants <-  bind_rows(pinterest_other, highfashions_other)
other_pants <- bind_rows(other_pants, nordstrom_other)
other_pants$Trouser <- "no" # make new row for boolean variable trousers and assign all to "no"

```

```{r}
# make data frame with all trousers
pinterest_trousers <- pinterest_pants %>% filter(Embellishment ==  "trousers")
highfashions_trousers <- highfashion_spring_pants %>% filter(embellishment ==  "pleated")
nordstrom_trousers <- nordstrom_pants %>% filter(Material == "linen/cotton")


trousers <-  bind_rows(pinterest_trousers, highfashions_trousers)
trousers <- bind_rows(trousers, nordstrom_trousers)
trousers$Trouser <- "yes" # make new row for boolean variable trousers and assign all to "yes"
```

```{r}
# combine both trouser and non-trouser data frames
all_pants <-  bind_rows(trousers, other_pants)
```

### Write hypotheses & check conditions:

Null Hypothesis: prop1 = prop2 Alternative Hypothesis: prop1 not equal
to prop2

We can assume that the data is approximately normal since n \> 30. The
data was randomly collected.

Set alpha at 0.05

```{r}
# calculate sample size and proportion
sample_size = 40
sample_prop = (14/40)
hypoth_prop = 0.5 # where proportion of trousers are equal to other types of pants
```

```{r}
# perform z test
test <- prop.test( x = sample_prop, n = sample_size, p = hypoth_prop )
test
```
Since the p-value is less than alpha = 0.05, there is enough evidence to
reject the null hypothesis. This means that the proportion of trousers
is not equal to the proportion of other pants. Based on this test, we
can conclude that trousers might be overtaking the other styles of pants in
popularity among various retailers.

Visualization:
```{r}
ggplot(all_pants, aes(x = Trouser)) +
  geom_bar(fill = "lightpink") + 
  labs(title = "Trousers vs. Non-Trousers",
       x = "Trousers?",
       y = "Count")
```

### Trousers Conclusion:

We were specifically curious about investigating this because it is a trend that everyone on our team has witnessed in our day-to-day life. That is one of the special things about the way we conducted our research: we got to investigate whatever we wanted to. We had a suspicion that trousers are the most popular type of pants for women at the moment, and our data reaffirmed our hypothesis

### From Our Team:

We have thoroughly enjoyed exploring our interests in fashion and data science throughout this research project, and we have collectively learned so much. Although still at an elementary level, this project was an opportunity to follow-through with every step of data science research, from data collection to forming conclusions. We would be interested in exploring these topics further in the future, and as our skills develop, we will be able to discover even more interesting results.

Thank you for your interest in our work,

The Quetzal Team:

Esha Chakrabarty, Hannah Wen, Jamie Chac, Libby Amir, & Pearl Vishen   
